Modeling the impacts of replacing car trips with combined public transport and cycling: \ Reproducible methods, results and actionable evidence from biclaR

### 2024-05-03 paper/PaperCEUS/PaperCEUS_rev1.Rmd
solve bib

---
title: "Modeling the impacts of replacing car trips with combined public transport and cycling: \ Reproducible methods, results and actionable evidence from biclaR"
author:
  - name: Rosa Félix\corref{cor1}
    affiliation: CERIS
    correspondingauthor: true
    # footnote: 1
    email: rosamfelix@tecnico.ulisboa.pt
  - name: Filipe Moura
    affiliation: CERIS
    email: fmoura@tecnico.ulisboa.pt
  - name: Robin Lovelace
    affiliation: ITS
    email: r.lovelace@uleeds.uk
address:
  - code: CERIS
    address: "CERIS, Instituto Superior Técnico - Universidade de Lisboa, Av Rovisco Pais 1, 1049-001 Lisboa, Portugal"
  - code: ITS
    address: "Institute for Transport Studies, University of Leeds. 34-40 University Rd, Leeds LS2 9JT, UK"
abstract: |
  A high proportion of car trips can be replaced by a combination of public transit and cycling for the first-and-last mile. This paper estimates the potential for cycling combined with public transit as a substitute for car trips in the Lisbon metropolitan area and assesses its socio-environmental impacts using open data and open source tools.
  A decision support tool that facilitates the design and development of a metropolitan cycling network was developed (_biclaR_). The social and environmental impacts were assessed using the _HEAT for Cycling_ and the _HEAT as a Service_ tools. The impacts of shifting car trips to public transport were also estimated and monetized.
  The results show that 10% of all trips could be made by cycling in combination with public transport. Shifting to cycling for the shorter first and last mile stages can reduce annual CO~2~eq emissions from 3,000 to 7,500 tons/day, while for the public transport leg, the transfer from car avoids of up to 20,500 tons of CO~2~eq emissions per year. The estimated socio-environmental benefits are of €125 million to €325 million over 10 years.
  This evidence can support policymakers to prioritize interventions that reduce the reliance on private motor vehicles.
  \
keywords: 
  - Active transport
  - Intermodality
  - First and last mile
  - Health economic assessment
  - Environmental impacts
  - Open data and methods
journal: "Computers, Environment and Urban Studies"
date: "`r Sys.Date()`"
classoption: review, doubleblind, 3p, authoryear #change later to preprint instead of review
bibliography: bibliography.bib
biblio-style: elsarticle-harv
linenumbers: true
numbersections: true
# Use a CSL with `citation_package = "default"`
csl: https://www.zotero.org/styles/computers-environment-and-urban-systems
# csl: computers-environment-and-urban-systems.csl
# csl: trd_test.csl
preamble: |
  \biboptions{authoryear}
  \usepackage{url}
  \usepackage{float}
  \usepackage{booktabs}
  \usepackage{longtable}
  \usepackage{makecell}
  \usepackage{multirow}
output: 
  rticles::elsevier_article:
    keep_tex: yes
    citation_package: "default" #ou nenhum
    latex_engine: xelatex
    extra_dependencies: "subfig"
#   bookdown::word_document2:
#     fig_caption: true
# always_allow_html: true
urlcolor: blue
citecolor: blue
linkcolor: blue
# natbiboptions: longnamesfirst,angle,semicolon
---

```{r eval=FALSE, echo=FALSE}
system("latexdiff paper/PaperCEUS.tex paper/PaperCEUS_rev1.tex > paper/PaperCEUS_diff.tex")

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      # root.dir = "paper/PaperTRA/",
                      out.extra = "")  # This forces knitr to label all figures.

@@

```{r jitter, fig.align='center', fig.cap="Representation of a flow-weighted random sample of 10,000 desire lines in the Lisbon metropolitan area between districts, without jittering (left) and with jittering (right).", out.width="100%"}

# od_all = readRDS("paper/paperTRA/load/od_all.Rds")
# od_jittered_filter = readRDS("paper/paperTRA/load/od_jittered_filter.Rds")
# 
# par(mfrow = c(1, 2))
# plot(od_all$geometry, lwd = 0.1,)
# plot(od_jittered_filter$geometry, lwd = 0.1)
# knitr::include_graphics("img/jitter.png", error = FALSE)
# 
# 
# odcombined = rbind(
#   od_all |> 
#     transmute(Total, Disaggregation = " Original"),
#   od_jittered_filter |> 
#     transmute(Total, Disaggregation = "Jittered")
# )
# odcombined |> 
#   sample_n(10000, weight = odcombined$Total) |>
#   mutate(Total = Total / 1000) |> 
#   ggplot() +
#   geom_sf(aes(alpha = Total)) +
#   facet_wrap(~Disaggregation) +
#   theme_void() +
#   geom_sf(data = MUNICIPIOSgeo, fill = NA, color = "gray60")

knitr::include_graphics("img/jitter_rev.png", error = FALSE)
```

@@

Figure \ref{fig:map2} illustrates the resulting bicycle routes to access the main PT interfaces in the LMA.

```{r map2, out.width="80%", fig.cap="Bike routes with the highest potential to serve as first and last leg when replacing cycling and PT from car trips (screenshot of the interactive online tool).", fig.align='center'}
knitr::include_graphics("img/map2.png", error = FALSE)
```

>>>>

---
title: "Modeling the impacts of replacing car trips with combined public transport and cycling: \ Reproducible methods, results and actionable evidence from biclaR"
author:
  - name: Rosa Félix\corref{cor1}
    affiliation: CERIS
    correspondingauthor: true
    # footnote: 1
    email: rosamfelix@tecnico.ulisboa.pt
  - name: Filipe Moura
    affiliation: CERIS
    email: fmoura@tecnico.ulisboa.pt
  - name: Robin Lovelace
    affiliation: ITS
    email: r.lovelace@uleeds.uk
address:
  - code: CERIS
    address: "CERIS, Instituto Superior Técnico - Universidade de Lisboa, Av Rovisco Pais 1, 1049-001 Lisboa, Portugal"
  - code: ITS
    address: "Institute for Transport Studies, University of Leeds. 34-40 University Rd, Leeds LS2 9JT, UK"
abstract: |
  A high proportion of car trips can be replaced by a combination of public transit and cycling for the first-and-last mile. This paper estimates the potential for cycling combined with public transit as a substitute for car trips in the Lisbon metropolitan area and assesses its socio-environmental impacts using open data and open source tools.
  A decision support tool that facilitates the design and development of a metropolitan cycling network was developed (_biclaR_). The social and environmental impacts were assessed using the _HEAT for Cycling_ and the _HEAT as a Service_ tools. The impacts of shifting car trips to public transport were also estimated and monetized.
  The results show that 10% of all trips could be made by cycling in combination with public transport. Shifting to cycling for the shorter first and last mile stages can reduce annual CO~2~eq emissions from 3,000 to 7,500 tons/day, while for the public transport leg, the transfer from car avoids of up to 20,500 tons of CO~2~eq emissions per year. The estimated socio-environmental benefits are of €125 million to €325 million over 10 years.
  This evidence can support policymakers to prioritize interventions that reduce the reliance on private motor vehicles.
  \
keywords: 
  - Active transport
  - Intermodality
  - First and last mile
  - Health economic assessment
  - Environmental impacts
  - Open data and methods
journal: "Computers, Environment and Urban Studies"
date: "`r Sys.Date()`"
classoption: review, doubleblind, 3p, authoryear #change later to preprint instead of review
bibliography: bibliography.bib
biblio-style: elsarticle-harv
linenumbers: true
numbersections: true
# Use a CSL with `citation_package = "default"`
csl: https://www.zotero.org/styles/computers-environment-and-urban-systems
# csl: computers-environment-and-urban-systems.csl
preamble: |
  \biboptions{authoryear}
  \usepackage{url}
  \usepackage{float}
  \usepackage{booktabs}
  \usepackage{longtable}
  \usepackage{makecell}
  \usepackage{multirow}
output: 
  rticles::elsevier_article:
    keep_tex: yes
    # citation_package: "default" #or none
    latex_engine: xelatex
    extra_dependencies: "subfig"
#   bookdown::word_document2:
#     fig_caption: true
# always_allow_html: true
urlcolor: blue
citecolor: blue
linkcolor: blue
# natbiboptions: longnamesfirst,angle,semicolon
---

```{r eval=TRUE, echo=FALSE}
# setwd("paper/PaperCEUS/") # knit directory: Document
# system("latexdiff PaperCEUS.tex PaperCEUS_rev1.tex > tracked_changes_rev1.tex")
# system("pdflatex tracked_changes_rev1.tex")
```

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      # root.dir = "paper/PaperCEUS/",
                      out.extra = "")  # This forces knitr to label all figures.

@@

```{r jitter, fig.align='center', fig.cap="Flow-weighted random sample of 10,000 desire lines in the Lisbon metropolitan area between districts of the 18 municipalities, without jittering (left) and with jittering (right).", fig.height=4, out.width="100%"}

# od_all = readRDS("paper/paperTRA/load/od_all.Rds")
# od_jittered_filter = readRDS("paper/paperTRA/load/od_jittered_filter.Rds")
# 
# par(mfrow = c(1, 2))
# 
# odcombined = rbind(
#   od_all |>
#     transmute(Total, Disaggregation = " Original"),
#   od_jittered_filter |>
#     transmute(Total, Disaggregation = "Jittered")
# )
# odcombined |>
#   sample_n(10000, weight = odcombined$Total) |>
#   mutate(Total = Total / 1000) |>
#   ggplot() +
#   geom_sf(aes(alpha = Total)) +
#   facet_wrap(~Disaggregation) +
#   theme_void() +
#   theme(legend.position="none",
#         strip.text.x = element_blank()) +
#   geom_sf(data = MUNICIPIOSgeo, fill = NA, color = "gray60")
# 
# ggsave(filename = "img/jitter_rev2.png", plot = last_plot(), width = 1000, height = 400, units = "px", dpi = 300, type = "cairo")

knitr::include_graphics("img/jitter_new.png", error = FALSE)
```

@@

Figure \ref{fig:map2} illustrates the resulting bicycle routes to access the main PT interfaces in the LMA. It is noticeable the high potential of the train interfaces to attract car-to-PT substituting trips.

```{r map2, out.width="80%", fig.cap="Bike routes with the highest potential to serve as first and last leg when replacing cycling and PT from car trips (screenshot of the interactive online tool). Larger the line width, higher the segment's car to bicycle shift potential. Darker the line color, lower the quietness level.", fig.align='center'}
knitr::include_graphics("img/map2.png", error = FALSE)
```
